Paper_Figure

### 2024-03-26 paper_figure.Rmd
24/03/26

# Check Gene Diffusion on Cell graph by time
pl = lapply(1:length(genes),function(i){
  gene = genes[i]
  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(2)[i])
})
p = as.ggplot(pl[[1]]) / as.ggplot(pl[[2]] & labs(title = NULL))
ggsave(filename = "fig-pipeline3.png", plot = p, width = 16, height = 8)
```

@@

# Set the color code for all methods
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs"),method_ls)
coldef = setNames(
  colorRampPalette(brewer.pal(9, "Set1"))(length(method_fullname)),
  method_fullname)
```
```{r}
path = "./benchmark_result/"
tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(paste0(path,tissue_name,"_auroc.csv"),header = TRUE); df$'Tissue' = tissue_name; df$'Method' = method_fullname[df$'Method']; df}))

df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HighlyVariableGenes")
df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
levels(df_auc$gt_set) = c("Criterion1 - CellMarkerDB","Criterion2 - MaxFC")
df_auc$Tissue = factor(df_auc$Tissue,levels = tissue_ls)
levels(df_auc$Tissue) = c("Bone Marrow","Pancreas","Lung")
# Create the plot
fig <- ggplot(df_auc, aes(x = Tissue, y = AUROC, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
  labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method]) + facet_wrap(~gt_set)
fig = fig + theme(
          legend.title = element_text(size = 10),
          legend.text = element_text(size = 10),
          axis.title.x = element_text(size = 15),
          axis.title.y = element_text(size = 15),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 10),
          panel.grid = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.background = element_blank(),
          strip.background = element_rect(fill = "white", color = "white") ,
          strip.text = element_text(size = 15, color = "black"))

@@

# Fig - Density Index
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
score_measure = read.table(paste0(path,tissue_name,"_density_index.csv"),header = TRUE)
score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)

@@

score_measure$'color' = coldef[score_measure$Method]
score_measure[score_measure$Method == "All genes","color"] = "black"
score_measure$'line_type' = "solid"
score_measure[score_measure$Method == "All genes","line_type"] = "dashed"

score_measure[grepl("Seurat",score_measure$Method),"Method"] = "Seurat v4"

@@

# # Add null distribution
# null_score_measure = read.csv(paste0(path,tissue_name,"_DI_null.csv"),header = TRUE)
# fig = fig + geom_boxplot(data = null_score_measure, aes(x = TopGene, y = DensityIndex, group = TopGene), inherit.aes = FALSE)

ggsave(filename = "fig-densityindex.png", plot = fig, width = 5, height = 5)
```

>>>>

# Check Gene Diffusion on Cell graph by time
pl = lapply(1:length(genes),function(i){
  gene = genes[i]
  FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,W = W,check_time = check_time,gene_name = NULL, gene_color = scales::hue_pal()(length(genes))[i])
})
p = as.ggplot(pl[[1]]) / as.ggplot(pl[[2]] & labs(title = NULL))
ggsave(filename = "fig-pipeline3.png", plot = p, width = 16, height = 8)
```

@@

# Set the color code for all methods
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg","gene_count")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4 default", "Seurat v4 Wilcoxon unfiltered","HVGs","Gene count"),method_ls)
coldef = setNames(
  colorRampPalette(brewer.pal(9, "Set1"))(length(method_fullname)),
  method_fullname)
```
```{r}
folder_path = "./benchmark_result/"
tissue_ls = c("marrow_facs","pancreas_facs","lung_facs")
df_auc = do.call(rbind,lapply(tissue_ls,function(tissue_name){df = read.table(file.path(folder_path,paste0(tissue_name,"_auroc.csv")),header = TRUE); df$'Tissue' = tissue_name;df}))
df_auc$Method = method_fullname[df_auc$Method]
df_auc = df_auc %>% filter(gt_set %in% c("Top100","CellMarkerDB")) %>% filter(!Method %in% "HVGs")
df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
levels(df_auc$gt_set) = c("Criterion1 - CellMarkerDB","Criterion2 - MaxFC")
df_auc$Tissue = factor(df_auc$Tissue,levels = tissue_ls)
levels(df_auc$Tissue) = c("Bone Marrow","Pancreas","Lung")
# Create the plot
fig <- ggplot(df_auc, aes(x = Tissue, y = AUROC, fill = Method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = round(AUROC, 2)), position = position_dodge(width = 1), vjust = -0.25, size = 2.5, angle = 45) +
  labs(x = "Tissue", y = "AUROC", fill = "Method") + coord_cartesian(ylim = c(0.5, 1)) +
  theme(legend.position = "bottom")  +  scale_fill_manual(values= coldef[names(coldef)%in%df_auc$Method]) + facet_wrap(~gt_set)
fig = fig + theme(
          legend.title = element_text(size = 10),
          legend.text = element_text(size = 10),
          axis.title.x = element_text(size = 15),
          axis.title.y = element_text(size = 15),
          axis.text.x = element_text(size = 12),
          axis.text.y = element_text(size = 10),
          panel.grid = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.background = element_blank(),
          strip.background = element_rect(fill = "white", color = "white") ,
          strip.text = element_text(size = 15, color = "black"))

@@

# Fig - Density Index
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
score_measure = read.table(paste0(path,tissue_name,"_density_index_no_dropout.csv"),header = TRUE)
score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)

@@

score_measure$'color' = coldef[score_measure$Method]
score_measure[score_measure$Method == "All genes","color"] = "black"
score_measure$'line_type' = "solid"
score_measure[score_measure$Method == "All genes","line_type"] = "dashed"
score_measure[grepl("Seurat",score_measure$Method),"Method"] = "Seurat v4"

@@

# Add null distribution
null_score_measure = read.csv(paste0(path,tissue_name,"_DI_null.csv"),header = TRUE)
fig = fig + geom_boxplot(data = null_score_measure, aes(x = TopGene, y = DensityIndex, group = TopGene), inherit.aes = FALSE)

ggsave(filename = "fig-densityindex.png", plot = fig, width = 5, height = 5)
```
## Scatterplot
```{r}
score_measure = read.table(paste0(path,tissue_name,"_density_index_no_dropout.csv"),header = TRUE)
score_measure = score_measure %>% filter(Method!="wilcox_no_filter")
score_measure$Method = ifelse(score_measure$Method %in% names(method_fullname),method_fullname[score_measure$Method],score_measure$Method)
score_measure$'color' = coldef[score_measure$Method]

ggplot(score_measure %>% filter(Method!="All genes") %>% filter(TopGene <= 1000), aes(x = non_dropout_rate, y = Mean_SI, color = Method)) + 
  geom_point() +
  scale_color_manual(values = score_measure %>% filter(Method!="All genes") %>% select(c("Method","color")) %>% distinct() %>% tibble::deframe()) + facet_wrap(~TopGene)
```
## Prop of recovered cts
```{r}
recovery_rate <- reshape2::melt(recovery_rate,id.vars = "TopGene",value.name = "prop",variable.name = "Method")
recovery_rate = recovery_rate %>% filter(Method!="wilcox_no_filter")
recovery_rate$Method = ifelse(recovery_rate$Method %in% names(method_fullname),method_fullname[recovery_rate$Method],recovery_rate$Method)
recovery_rate$'color' = coldef[recovery_rate$Method]

ggplot(recovery_rate, aes(x = TopGene, y = prop, color = Method, group = Method)) + 
  geom_line() +
  scale_color_manual(values = score_measure %>% filter(Method!="All genes") %>% select(c("Method","color")) %>% distinct() %>% tibble::deframe()) + ylab("Prop. of recovered cell types")
```
